The relative roles of genetic, ecological and environmental factors in insular species radiations

### 2018-07-03 paper/draft.tex
include additional thoughts and refs in intro

%% Cite & look at Katja Schiffer paper
%% "species rich vs. species poor lineage''
%% ``emerging systematic, functional, phylogenetic patterns''
%% Hurlbert

%% major topic: besides classical mechanisms (ecological, environmental), genetic mechanisms affect speciation/radiations
Island radiations like the Darwin finches on Galápagos or \textit{Aeonium} on the Canaries have fascinated researchers throughout history %% cite? Darwin? ...?
To this date, however, it is unclear why these genera produced this multitude of species as opposed to most other genera that remained species poor.
Until recently, researchers focused on environmental and ecological factors to explain different speeds and ocurrences of speciation.\cite{krugSpeciesGenusRatios2008}\cite{hurlbertProcessesGeneratingLatitudinal2014}\cite{willisSpeciesDiversityScale2002}
%% HERE COME SOME EXAMPLES
Yet, these mechanisms often fail to explain how the rapid establishment of reproductive barriers or isolation arises,
which ultimately results in sympatric speciation. %CLAIM! PROVIDE REFERENCES!
%% somewhere here Patino et al 2018 anagenesis, cladogenesis, etc islands
With the growing importance of genetic data and analyses, an increasing number of studies also investigates genetic processes and
their effect on population variability. %EXAMPLES
Some of these studies highlight the importance of genetic linkage or ``hitchhiking'' on reproductive/ecological isolation of populations \cite{hawthorneGeneticLinkageEcological2001}\cite{viaGeneticMosaicSuggests2008}%(e.g., Via \& West 2008, Via 2001 and references therein)
as well as evolutionary rescue provided by the feedback with ecological processes \cite{schiffersLimitedEvolutionaryRescue2012}.
%%TODO: here come a list of processes with examples
%% - genetic processes in general
%% - isolation processes
Besides arthropods \cite{hawthorneGeneticLinkageEcological2001}, these processes can be observed in plants as well. %Are you talking about islands here? Be clearer, because you said about the lack of geentic studies in island contexts.
For example, the rapid switching of pollinator syndromes in species of the genus \textit{Petunia} %use package for this!
seems to be a consequence of the tight genetic linkage between loci for pollination syndrome traits \cite{hermannTightGeneticLinkage2013}
In an island context, genetic studies are even rarer and almost exclusively of a descriptive nature.

>>>>

%% Cite & look at Katja Schiffer paper
%% "species rich vs. species poor lineage''
%% ``emerging systematic, functional, phylogenetic patterns''
%% Hurlbert
% !! Hypothesis: Quick establishment of reproductive barriers (+ population variability) facilitates/speeds up adaptation. !!

%% major topic: besides classical mechanisms (ecological, environmental), genetic mechanisms affect speciation/radiations
Island radiations like the Darwin finches on Galápagos or \textit{Aeonium} on the Canaries have fascinated researchers throughout history %% cite? Darwin? ...?
However, it is often unclear why these genera produced this multitude of species as opposed to most other genera that remained species poor.
To explain speed and occurences of radiations, researchers still mainly focus on environmental and ecological factors \cite{krugSpeciesGenusRatios2008}\cite{hurlbertProcessesGeneratingLatitudinal2014}\cite{willisSpeciesDiversityScale2002}. % when radiations DID happen!
While these factors often correlate strongly with genetic divergence of species and populations, % REFS!
in some cases the reason for maintenance of population diversity remains in the dark. \cite{chenConcordanceGeneticDiversity}
%% HERE COME SOME EXAMPLES
Yet, these mechanisms often fail to explain how the rapid establishment of reproductive barriers or isolation arises,
which ultimately results in sympatric speciation. %CLAIM! PROVIDE REFERENCES!
%% somewhere here Patino et al 2018 anagenesis, cladogenesis, etc islands
With the growing importance of genetic data and analyses, an increasing number of studies also investigates genetic processes and
their effect on population variability. %EXAMPLES
Some of these studies highlight the importance of genetic linkage or ``hitchhiking'' on reproductive/ecological isolation of populations \cite{hawthorneGeneticLinkageEcological2001}\cite{viaGeneticMosaicSuggests2008}%(e.g., Via \& West 2008, Via 2001 and references therein)
as well as evolutionary rescue provided by the feedback with ecological processes \cite{schiffersLimitedEvolutionaryRescue2012}.
%%TODO: here come a list of processes with examples
%% - genetic processes in general
%% - isolation processes
Besides arthropods \cite{hawthorneGeneticLinkageEcological2001}, these processes can be observed in plants as well. %Are you talking about islands here? Be clearer, because you said about the lack of geentic studies in island contexts.
For example, the rapid switching of pollinator syndromes in species of the genus \textit{Petunia} %use package for this!
seems to be a consequence of the tight genetic linkage between loci for pollination syndrome traits \cite{hermannTightGeneticLinkage2013}
In an island context, genetic studies are even rarer and almost exclusively of a descriptive nature.
